The Changing Shape of Empire: A GM Study of Chimú Bottles in Museum Collections

### 2022-01-27 supp/01-linear.Rmd
<edit ms>

## Load data + define variables

```{r setwd ld define variables 3, echo=TRUE}
data <- read.csv("potmeas.csv", header = TRUE, as.is = TRUE)

# define variables
od <- data$orifice.diameter # orifice diameter
bd <- data$base.diameter # base diameter
nd <- data$bneck.diameter # neck diameter
bl <- data$wide.diameter # belly diameter
h <- data$height # height
rt <- data$rim.thickness # rim thickness
gp <- data$form # bottle or jar
```

@@

```{r boxsite, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Boxplots of orifice diameter, base diameter, neck diameter, belly diameter, height, and rim thickness for the anthropomorphic, phytomorphic, and zoomorphic Chimú vessels used in this study."}
# boxplot of orifice diameter ~ group
odgroup <- ggplot(data, aes(x = gp, y = od, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Orifice Diam") +
  theme(legend.position = "none")

# boxplot of base diameter ~ group
bdgroup <- ggplot(data, aes(x = gp, y = bd, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Base Diam") +
  theme(legend.position = "none")

# boxplot of neck diameter ~ group
ndgroup <- ggplot(data, aes(x = gp, y = nd, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Neck Diam") +  
  theme(legend.position = "none")

# boxplot of belly diameter ~ group
blgroup <- ggplot(data, aes(x = gp, y = bl, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Belly Diam") +  
  theme(legend.position = "none")

# boxplot of height ~ group
hgroup <- ggplot(data, aes(x = gp, y = h, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Height") +  
  theme(legend.position = "none")

# boxplot of rim thickness ~ group
rtgroup <- ggplot(data, aes(x = gp, y = rt, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Rim Thick") +  
  theme(legend.position = "none")

@@

```{r pca, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of Principal Components Analysis for linear metrics associated with the Chimu vessels used in this study (gray squares, anthropomorphic; orange triangles, phytomorphic; tan circles, zoomorphic)."}
# set plot parameters to plot by group
df <- data[c(3:8)]
pch.gps.gp <- c(15,17)[as.factor(gp)]
col.gps.gp <- wes_palette("Moonrise2")[as.factor(gp)]

## pca plot
pca <- autoplot(prcomp(df),
                data = data,
                asp = 1,
                shape = pch.gps.gp,
                colour = "form",
                variance_percentage = TRUE,
                loadings = TRUE, 
                loadings.colour = 'blue',
                loadings.label = TRUE,
                loadings.label.size = 3,
                frame = TRUE,
                frame.type = 't') +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2"))

@@

## Density plots for _variable_ ~ _group_

### _Orifice diameter_ ~ _group_

```{r density.gp.od, out.width = "100%", dpi = 300, echo=TRUE, fig.cap="Density plot for orifice diameter by group."}
# density plot for orifice diameter by group
Group <- data$form
ggplot(data, aes(x = od, fill = Group, colour = Group)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  xlim(0,10) +
  labs(x = 'Orifice Diameter', y = 'density')
```

### _Base diameter_ ~ _group_

```{r density.gp.bd, out.width = "100%", dpi = 300, echo=TRUE, fig.cap="Density plot for base diameter by group."}
# density plot for base diameter by group
ggplot(data, aes(x = bd, fill = Group, colour = Group)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  xlim(0,7) +
  labs(x = 'Base Diameter', y = 'density')
```

### _Neck diameter_ ~ _group_

```{r density.gp.nd, out.width = "100%", dpi = 300, echo=TRUE, fig.cap="Density plot for neck diameter by group."}
# density plot for neck diameter by group
ggplot(data, aes(x = nd, fill = Group, colour = Group)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  xlim(0,10) +
  labs(x = 'Neck Diameter', y = 'density')
```

### _Belly diameter_ ~ _group_

```{r density.gp.bl, out.width = "100%", dpi = 300, echo=TRUE, fig.cap="Density plot for belly diameter by group."}
# density plot for belly diameter by group
ggplot(data, aes(x = bl, fill = Group, colour = Group)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  xlim(5,33) +
  labs(x = 'Belly Diameter', y = 'density')
```

### _Height_ ~ _group_

```{r density.gp.h, out.width = "100%", dpi = 300, echo=TRUE, fig.cap="Density plot for vessel height by group."}
# density plot for height by group
ggplot(data, aes(x = h, fill = Group, colour = Group)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  xlim(8,30) +
  labs(x = 'Vessel Height', y = 'density')
```

@@

```{r anova.gp.od, out.width = "100%", dpi = 300, echo=TRUE}
# anova = orifice diameter ~ group
gp.od <- lm.rrpp(od ~ gp, 
                 SS.type = "I", 
                 data = data, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.od)
```

@@

```{r anova.gp.bd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = base diameter ~ group
gp.bd <- lm.rrpp(bd ~ gp, 
                 SS.type = "I", 
                 data = data, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bd)
```

@@

```{r anova.gp.nd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = neck diameter ~ group
gp.nd <- lm.rrpp(nd ~ gp, 
                 SS.type = "I", 
                 data = data, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.nd)
```

@@

```{r anova.gp.bl, out.width = "100%", dpi = 300, echo=TRUE}
# anova = belly diameter ~ group
gp.bl <- lm.rrpp(bl ~ gp, 
                 SS.type = "I", 
                 data = data, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bl)
```

@@

```{r anova.gp.h, out.width = "100%", dpi = 300, echo=TRUE}
# anova = height ~ group
gp.h <- lm.rrpp(h ~ gp, 
                SS.type = "I", 
                data = data, 
                iter = 9999, 
                print.progress = FALSE)
anova(gp.h)
```

>>>>

## Load DATA + define variables

```{r setwd ld define variables 3, echo=TRUE}
DATA <- read.csv("potmeas.csv", header = TRUE, as.is = TRUE)

# define variables
od <- DATA$od <- DATA$orifice.diameter # orifice diameter
bd <- DATA$bd <- DATA$base.diameter # base diameter
nd <- DATA$nd <- DATA$bneck.diameter # neck diameter
bl <- DATA$bl <- DATA$wide.diameter # belly diameter
h <- DATA$h <- DATA$height # height
RT <- DATA$RT <- DATA$rim.thickness # rim thickness
gp <- DATA$gp <- as.factor(DATA$form) # bottle or jar
```

@@

```{r boxsite, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Boxplots of orifice diameter, base diameter, neck diameter, belly diameter, height, and rim thickness for the anthropomorphic, phytomorphic, and zoomorphic Chimú vessels used in this study."}
# boxplot of orifice diameter ~ group
odgroup <- ggplot(DATA, aes(x = gp, y = od, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Orifice Diam") +
  theme(legend.position = "none")

# boxplot of base diameter ~ group
bdgroup <- ggplot(DATA, aes(x = gp, y = bd, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Base Diam") +
  theme(legend.position = "none")

# boxplot of neck diameter ~ group
ndgroup <- ggplot(DATA, aes(x = gp, y = nd, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Neck Diam") +  
  theme(legend.position = "none")

# boxplot of belly diameter ~ group
blgroup <- ggplot(DATA, aes(x = gp, y = bl, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Belly Diam") +  
  theme(legend.position = "none")

# boxplot of height ~ group
hgroup <- ggplot(DATA, aes(x = gp, y = h, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Height") +  
  theme(legend.position = "none")

# boxplot of rim thickness ~ group
rtgroup <- ggplot(DATA, aes(x = gp, y = RT, color = gp)) + 
  geom_boxplot() +
  geom_dotplot(binaxis = 'y',stackdir = 'center',dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Form", y = "Rim Thick") +  
  theme(legend.position = "none")

@@

```{r pca, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of Principal Components Analysis for linear metrics associated with the Chimu vessels used in this study (gray squares, anthropomorphic; orange triangles, phytomorphic; tan circles, zoomorphic)."}
# set plot parameters to plot by group
df <- DATA[c(3:8)]
pch.gps <- c(15,17)[as.factor(gp)]
col.gps <- wes_palette("Moonrise2")[as.factor(gp)]

## pca plot
pca <- autoplot(prcomp(df),
                data = DATA,
                asp = 1,
                shape = pch.gps,
                colour = "form",
                variance_percentage = TRUE,
                loadings = TRUE, 
                loadings.colour = 'blue',
                loadings.label = TRUE,
                loadings.label.size = 3,
                frame = TRUE,
                frame.type = 't') +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_colour_manual(values = wes_palette("Moonrise2"))

@@

@@

```{r anova.gp.od, out.width = "100%", dpi = 300, echo=TRUE}
# anova = orifice diameter ~ group
gp.od <- lm.rrpp(od ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.od)
```

@@

```{r anova.gp.bd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = base diameter ~ group
gp.bd <- lm.rrpp(bd ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bd)
```

@@

```{r anova.gp.nd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = neck diameter ~ group
gp.nd <- lm.rrpp(nd ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.nd)
```

@@

```{r anova.gp.bl, out.width = "100%", dpi = 300, echo=TRUE}
# anova = belly diameter ~ group
gp.bl <- lm.rrpp(bl ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bl)
```

@@

```{r anova.gp.h, out.width = "100%", dpi = 300, echo=TRUE}
# anova = height ~ group
gp.h <- lm.rrpp(h ~ gp, 
                SS.type = "I", 
                DATA = DATA, 
                iter = 9999, 
                print.progress = FALSE)
anova(gp.h)
```

### ANOVA _Rim Thickness_ of Chimú vessels

```{r anova.gp.rt, out.width = "100%", dpi = 300, echo=TRUE}
# anova = rim thickness ~ group
gp.rt <- lm.rrpp(RT ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.rt)
```

### 2022-01-27 supp/index.Rmd
<edit ms>

 * Bottles and jars **differ significantly** in orifice diameter
 * Bottles and jars **differ significantly** in belly diameter

>>>>

 * Bottles and jars **differ significantly** in orifice diameter
 * Bottles and jars **differ significantly** in belly diameter
 * Bottles and jars **differ significantly** in rim thickness
